---
title: "Health Survey Data"
format: html
editor: visual
---

```{r}
# Load required packages
library(readxl)
library(dplyr)
```

## Creating Data Dictionary

```{r, warning = FALSE}
# Create a list of the titles of all sheets in workbook
sheets <- excel_sheets("~/Documents/GitHub/stat3888/AHSinpBasicCURFdataItemList-2.xls")

# Initialise data frame to store dictionary
combined_dictionary = data.frame()

# Remove contents + detailed index sheets
# Remove appendix classification sheets
sheets <- sheets[c(-1,-2, -18, -19, -22, -23, -24, -25, -26, -27)] 

# Loop through each of the selected sheets
for (sheet in sheets){
  # Display in console which sheet we are up to in the loop
  message(sheet)
  
  # Read the sheet into a dataframe
  df <- read_excel("~/Documents/GitHub/stat3888/AHSinpBasicCURFdataItemList-2.xls",
                   sheet = sheet, skip = 6)
  
  # Remove the empty rows and keep only useful columns
  df <- df %>%
    filter(!is.na(`Items & categories`)) %>%
    dplyr::select(`SAS name`, `Items & categories`, `Main population`)

  # Stack the sheet's rows onto the bottom of the combined dictionary
  combined_dictionary <- bind_rows(combined_dictionary, df)
}

```

## Investigating csv file contents

The BSP file seems to contain a large number (if not all) of the useful variables. Thus, we will clean only the data from this file.

```{r}
# Create list of all csv file paths
path <- "~/Desktop/Education/2025/STAT3888/Indigenous Dataset"
file_list <- list.files(path = path, pattern = "\\.csv?$", full.names = TRUE)

# Read files and store in list
df_list <- lapply(file_list, read.csv)

# Use the name of the file as the list index
names(df_list) <- basename(file_list)

# Filter the dictionary to include only the fields from the BSP file
inp13bad <- filter(combined_dictionary, 
                     `SAS name` %in% colnames(df_list$`inp13bad (1).csv`))
View(inp13bad)
                   
inp13bbi <- filter(combined_dictionary, 
                     `SAS name` %in% colnames(df_list$`inp13bbi.csv`))
View(inp13bbi)

inp13bcd <- filter(combined_dictionary, 
                     `SAS name` %in% colnames(df_list$`inp13bcd.csv`))
View(inp13bcd)

inp13bcn <- filter(combined_dictionary, 
                     `SAS name` %in% colnames(df_list$`inp13bcn.csv`))

View(inp13bcn)

inp13bcp <- filter(combined_dictionary, 
                     `SAS name` %in% colnames(df_list$`inp13bcp.csv`))

View(inp13bcp)

inp13bfd <- filter(combined_dictionary, 
                     `SAS name` %in% colnames(df_list$`inp13bfd.csv`))
View(inp13bfd)

inp13bhh <- filter(combined_dictionary, 
                     `SAS name` %in% colnames(df_list$`inp13bhh.csv`))
View(inp13bhh)

inp13bpa <- filter(combined_dictionary, 
                     `SAS name` %in% colnames(df_list$`inp13bpa.csv`))
View(inp13bpa)

inp13bpe <- filter(combined_dictionary, 
                     `SAS name` %in% colnames(df_list$`inp13bpe.csv`))
View(inp13bpe)

inp13bsp <- filter(combined_dictionary, 
                     `SAS name` %in% colnames(df_list$`inp13bsp.csv`))
View(inp13bsp)

inp13bsu <- filter(combined_dictionary, 
                     `SAS name` %in% colnames(df_list$`inp13bsu.csv`))
View(inp13bsu)

inp13bta <- filter(combined_dictionary, 
                     `SAS name` %in% colnames(df_list$`inp13bta.csv`))
View(inp13bta)
                   
```

```{r}

ggplot(data = df_list[["inp13bhh.csv"]], aes(x = factor(RAECURF), fill = factor(RAECURF))) +
  scale_x_discrete(labels = c("1" = "Non-remote", "2" = "Remote")) +
  geom_bar() +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(
    title = "Frequency of Remote and Non-Remote Survey Participants",
    x = element_blank(),
    y = "Frequency",
  )

ggplot(data = df_list[["inp13bsp.csv"]], aes(x = factor(ABSBID), fill = factor(ABSBID))) +
  geom_bar() +
  theme_minimal()

View(df_list$inp13bsp.csv)
```

```{r}
filtered_bsp <- df_list$inp13bsp.csv %>%
  dplyr::select(ABSGID, ABSHID, SMOKQ2AE, ALCPER1, SUGPER1, HEDATTC)

colnames(filtered_bbi) <- c("Person ID", "Household ID", "Smoking Status")

filtered_bbi <- df_list$inp13bbi.csv %>%
  dplyr::select(ABSGID, ABSHID, DIABPRVE, DIABFPGE, DIABQ01A)


View(filtered_data)

```

Summary of csv file contents
- inp13bad: Food group intake
- inp13bbi: Blood test results? (Cholestrol, glutamyl transferase)
- inp13bcd: Physical activity
- inp13bhh: Household information
- inp13bsb: EVERYTHING!!


VARIABLES NEEDED

- Diabetes [BBI]

DIABPRVE
Diabetes prevalence - HbA1c (%)
Persons aged 18 years and over who participated in the biomedical component

DIABFPGE
Diabetes prevalence - fasting plasma glucose (mmol/L)
Persons aged 18 years and over who participated in the biomedical component and fasted for 8 hours or more

- Family History of Diabetes [BSP]

DIABQ01A
Family history of Diabetes
All persons

- Remoteness [BHH]

RAECURF
Remoteness Area (ASGS 2011) [BHH]
Households


- Smoking status [BSP]

SMKPAROX
Presence of parent during smoking questions
All selected persons 15–17 years

SMKSTAT
Smoker status
All persons aged 15 years and over

SMKDAILY
Daily smoker status
All persons aged 15 years and over

SMOKQ2AE
Number of cigarettes usually smokes daily
All persons aged 15 years and over


- Alcohol [BSP]

ALCPER1
Percentage of energy from alcohol (Day1)
All persons


- Sugar intake [BSP]

SUGPER1
Percentage of energy from total sugars (Day1)
All persons

- Education [BSP]
	
HEDATTC
Highest level of educational attainment
All persons aged 15 years and over



TWO MODELS:
Logistic regression: diabetes or not? One for each region. Look at the significance of each predictor variable.

Interperet the models to figure out which predictor variables are most significant in each region. 


